---
title: "Edge_Limma"
author: "Harry Dolan"
date: "17/09/2020"
output: html_document
---
## read in feature counts and set names and condition...
```{r}
files <- read.delim("Rsubread.txt")
names(files) = c("Control_2", "Control_3", "Guide1_2", "Guide1_3", "Guide2_2", "Guide2_3", "Control_1", "Guide1_1", "Guide2_1")
group <- c("WT", "WT", "MT", "MT", "MT", "MT", "WT", "MT", "MT")
samplenames <- c("Control_2", "Control_3", "Guide1_2", "Guide1_3", "Guide2_2", "Guide2_3", "Control_1", "Guide1_1", "Guide2_1")
cond <- c("Control", "Control", "Guide1", "Guide1", "Guide2", "Guide2", "Control", "Guide1", "Guide2")
set <- c("New", "New", "New", "New", "New", "New", "Old", "Old", "Old")

```


## DGE Object
```{r}
library(edgeR)
x <- DGEList(counts = as.matrix(files))
colnames(x) <- samplenames
x$samples$group <- group
x$samples$cond <- cond
x$samples$set <- set

x$samples
```

## CPM, logCPM, + Mean and Median Lib Size
```{r}
cpm <- cpm(x)
lcpm <- cpm(x, log=TRUE)

L <- mean(x$samples$lib.size) * 1e-6
M <- median(x$samples$lib.size) * 1e-6
c(L, M)
```

## How many genes have a zero count?
```{r}
table(rowSums(x$counts==0)==9)
```

## EdgeR Keep Exprs Cull
```{r}
library(edgeR)
keep.exprs <- filterByExpr(x, group=group)
x <- x[keep.exprs,, keep.lib.sizes=FALSE]
dim(x)
```

Log-CPM density Plot
```{r}
library(RColorBrewer)

lcpm.cutoff <- log2(10/M + 2/L)
nsamples <- ncol(x)
col <- brewer.pal(nsamples, "Paired")
par(mfrow=c(1,2))
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
title(main="A. Raw data", xlab="Log-cpm")
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend(0, 0.25, samplenames, text.col=col, bty="n")
lcpm <- cpm(x, log=TRUE)
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
title(main="B. Filtered data", xlab="Log-cpm")
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend(0, 0.25, samplenames, text.col=col, bty="n")
```

## Normalisation Factors + Normalisation
```{r}
prenorm <- x
x <- calcNormFactors(x, method = "TMM")
x$samples$norm.factors

par(mfrow=c(1,2))
lcpm <- cpm(prenorm, log=TRUE)
boxplot(lcpm, las=2, col=col, main="")
title(main="A: Unnormalised data", ylab="Log-cpm")
x <- calcNormFactors(x)
x$samples$norm.factors

lcpm <- cpm(x, log=TRUE)
boxplot(lcpm, las=2, col=col, main="")
title(main="B: Normalised data", ylab="Log-cpm")
```

## PCA (LogCPM variation)
```{r}
col.group <- group

levels(col.group) <- brewer.pal(nlevels(col.group), "Set1")
col.group <- as.character(col.group)
col.cond <- cond
levels(col.cond) <- brewer.pal(nlevels(col.cond), "Set2")
col.cond <- as.character(col.cond)

plotMDS(lcpm, cex = 0.75, labels=(colnames(x)), col=c(rep("blue",6), rep("red",3)) )
title(main="A. Sample groups")

```

## Limma Model Matrix and Contrast Matrix
```{r}
design <- model.matrix(~0+cond+set)
colnames(design) <- gsub("group", "", colnames(design))
design

contr.matrix <- makeContrasts(
  Guide1vsControl = condGuide1 - condControl,
  Guide2vsControl = condGuide2 - condControl,
  Guide1vsGuide2 = condGuide1 - condGuide2,
  levels = colnames(design))
contr.matrix
```

## Limma Voom
```{r}
par(mfrow=c(2,2))
v <- voom(x, design, plot=TRUE)
v

vfit <- lmFit(v, design)
vfit <- contrasts.fit(vfit, contrasts=contr.matrix)
efit <- eBayes(vfit)
plotSA(efit)
title("voom: Mean-variance trend (post-fit)")
```
## diffeentially expressed genes
```{r}
tt_G1 <- topTable(efit, coef=1, n=18209)
tt_G2 <- topTable(efit, coef=2, n=18209)
G1_P <- tt_G1[tt_G1$adj.P.Val <= 0.05, ]
G2_P <- tt_G2[tt_G2$adj.P.Val <= 0.05, ]
G2_P_FC <- sum((tt_G1$adj.P.Val <= 0.05) & (tt_G1$logFC >= abs(1.5)))
G2_P_FC <- sum((tt_G2$adj.P.Val <= 0.05) & (tt_G2$logFC >= abs(1.5)))
```

##### G1 Volcanos ######
```{r}
library(EnhancedVolcano)
par(mfrow=c(2,2))
EnhancedVolcano(tt_G1,
                lab = rownames(tt_G1),
                x = 'logFC',
                y = 'P.Value',
                xlim = c(-5, 4),
                ylim = c(0, 6),
                title = 'Guide 1 CLN5 Knockdown vs Control',
                pCutoff = 0.00001,
                FCcutoff = 0,
                pointSize = 3.0,
                labSize = 1.0)
```

## GSEA (min size 20, max size 200)
```{r}
library(tidyverse)
tstats <- efit$t[,1]
tstats <- tstats %>% sort(., decreasing=TRUE)
library(ReactomePA)
rPAgsea = gsePathway(tstats, organism = "human", nPerm=10000, minGSSize = 20, pvalueCutoff = 0.05, maxGSSize = 200, pAdjustMethod = 'BH', verbose = FALSE
                     )
rPAgsea %>% as.data.frame() %>% View()
```

## GSEA2 (min size 20, no max size // parent brnaches)
```{r}
rPAgsea2 = gsePathway(tstats, organism = "human", nPerm=10000, minGSSize = 20, pvalueCutoff = 0.05, maxGSSize = 10000, pAdjustMethod = 'BH', verbose = FALSE
                     )
rPAgsea2 %>% as.data.frame() %>% View()
```

```{r}
```


```{r}
```


```{r}
```